---
title: "Replication: Beaudry and Portier (2014)"
subtitle: "News-Driven Business Cycles: Insights and Challenges"
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

## Overview

This document replicates the news shock identification of @beaudryportier2014. A news shock is identified as the shock that (i) has **no contemporaneous effect on TFP** and (ii) **maximizes the long-run forecast error variance of TFP** at a 40-quarter horizon. The approach follows the mixed restriction strategy combining a zero impact restriction with a long-run maximization criterion.

## Setup

```{r}
#| label: setup
pkgload::load_all(here::here())  # recompiles changed C++ files and loads package

library(tidyverse)
library(tictoc)

set_theme(ftheme_tidyMacro())
```

## Data

```{r}
#| label: data
load(here::here('data', 'BeaudryPortier2014.rda'))

# Variables: log TFP, log SP500, log Consumption, log Hours
y     <- BeaudryPortier2014 |> as.matrix()
T_obs <- nrow(y)
N     <- ncol(y)
```

## VAR Estimation

```{r}
#| label: var-estimation

p <- 2
c <- 1

# Estimate VAR: y_t = c + A_1 y_{t-1} + ... + A_p y_{t-p} + e_t
var_result <- fVAR(y, p = p, c = c)

# Variance-covariance matrix of reduced-form residuals
Sigma <- var_result$sigma_full
```

## News Shock Identification

The news shock is identified as the shock that has zero impact on TFP but maximises the cumulative long-run response of TFP over the 40-quarter horizon. Let $\Psi_h$ denote the $h$-step Wold coefficient matrices and $S$ the lower Cholesky factor of $\Sigma$. We search for a unit vector $\mathbf{h}$ such that the structural impact vector is:

$$
\mathbf{b} = S \begin{pmatrix} 0 \\ \mathbf{h}/\|\mathbf{h}\| \end{pmatrix}
$$

and the long-run effect on TFP, $\sum_{h=0}^{H} e_1^\top \Psi_h \mathbf{b}$, is maximized.

```{r}
#| label: identification

horizon <- 40

# Wold IRFs by inverting (I - A(L))
wold <- fwoldIRF(var_result, horizon = horizon)

# Lower Cholesky factor of Sigma
S <- t(chol(Sigma))

# Objective: minimise negative LR effect on variable 1 (TFP)
# subject to zero impact (first element of h2 fixed to 0)
lr_max <- function(theta, wold, S, var_idx) {
  h2        <- c(0, theta / sqrt(sum(theta^2)))
  lr_effect <- sum(sapply(1:dim(wold)[3], function(h) wold[var_idx, , h] %*% S %*% h2))
  return(-lr_effect)
}

set.seed(12345)
h0     <- rep(1, N - 1)   # N=4 variables, so 3 free parameters
result <- optim(
  par     = h0,
  fn      = lr_max,
  wold    = wold,
  S       = S,
  var_idx = 1,
  method  = "Nelder-Mead",
  control = list(reltol = 1e-10, maxit = 1e6)
)

h  <- result$par
h2 <- c(0, h / sqrt(sum(h^2)))

# Structural IRFs
struct_irf <- matrix(0, nrow = N, ncol = horizon + 1)
for (hh in 1:(horizon + 1)) {
  struct_irf[, hh] <- wold[,, hh] %*% S %*% h2
}
```

## Impulse Response Functions

```{r}
#| label: fig-irf-news
#| fig-cap: "Impulse response to the news shock"

varnames <- c("TFP", "SP500", "Consumption", "Hours")

irf_df <- as.data.frame(t(struct_irf))
colnames(irf_df) <- varnames
irf_df$horizon <- 0:horizon

irf_df |>
  pivot_longer(-horizon, names_to = "variable", values_to = "response") |>
  mutate(variable = factor(variable, levels = varnames)) |>
  ggplot(aes(x = horizon, y = response)) +
  geom_line() +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  facet_wrap(~variable, ncol = 2, scales = "free_y") +
  labs(x = "Horizon (quarters)", y = NULL)
```

## Bootstrap Confidence Bands: LR-Max Identification

```{r}
#| label: bootstrap-max

tic()
boot_max <- fbootstrapMaxCorrected(
  y          = y,
  var_result = var_result,
  nboot1     = 2000,
  nboot2     = 4000,
  horizon    = horizon,
  var_idx    = 1L,        # maximise long-run effect on TFP (variable 1)
  prc        = 68,
  cumulate   = integer(0) # no cumulation (levels data)
)
toc()
```

```{r}
#| label: fig-irf-boot-max
#| fig-cap: "IRF to news shock with 68% LR-Max bootstrap bands"

varnames <- c("TFP", "SP500", "Consumption", "Hours")

irf_to_df <- function(mat, label) {
  as.data.frame(t(mat)) |>
    setNames(varnames) |>
    mutate(horizon = 0:(ncol(mat) - 1), type = label)
}

bind_rows(
  irf_to_df(struct_irf,         "Point"),
  irf_to_df(boot_max$upper,     "Upper"),
  irf_to_df(boot_max$lower,     "Lower")
) |>
  pivot_longer(-c(horizon, type), names_to = "variable", values_to = "response") |>
  mutate(variable = factor(variable, levels = varnames)) |>
  pivot_wider(names_from = type, values_from = response) |>
  ggplot(aes(x = horizon)) +
  geom_ribbon(aes(ymin = Lower, ymax = Upper), alpha = 0.2) +
  geom_line(aes(y = Point)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  facet_wrap(~variable, ncol = 2, scales = "free_y") +
  labs(x = "Horizon (quarters)", y = NULL,
       caption = "Shaded region: 68% LR-Max bootstrap confidence band")
```

## Bootstrap Confidence Bands: Uhlig Max-Share Identification

```{r}
#| label: bootstrap-uhlig

tic()
boot_uhlig <- fbootstrapUhlig(
  y          = y,
  var_result = var_result,
  nboot      = 1000,
  horizon    = horizon,
  idx        = 1L,        # maximise FEV share of TFP (variable 1)
  prc        = 68,
  cumulate   = integer(0)
)
toc()
```

```{r}
#| label: fig-irf-boot-uhlig
#| fig-cap: "IRF to news shock with 68% Uhlig max-share bootstrap bands"

# Point estimate via Uhlig identification
h2_uhlig <- fuhlig_maxshare(wold, S, idx = 1L)
# Fix sign: last-horizon non-structural response of TFP should be positive
last_ns <- (wold[,, dim(wold)[3]] %*% S)
if (sum(last_ns[1, ] * h2_uhlig) < 0) h2_uhlig <- -h2_uhlig

struct_irf_uhlig <- matrix(0, nrow = N, ncol = horizon + 1)
for (hh in 1:(horizon + 1)) {
  struct_irf_uhlig[, hh] <- wold[,, hh] %*% S %*% h2_uhlig
}

bind_rows(
  irf_to_df(struct_irf_uhlig,   "Point"),
  irf_to_df(boot_uhlig$upper,   "Upper"),
  irf_to_df(boot_uhlig$lower,   "Lower")
) |>
  pivot_longer(-c(horizon, type), names_to = "variable", values_to = "response") |>
  mutate(variable = factor(variable, levels = varnames)) |>
  pivot_wider(names_from = type, values_from = response) |>
  ggplot(aes(x = horizon)) +
  geom_ribbon(aes(ymin = Lower, ymax = Upper), alpha = 0.2) +
  geom_line(aes(y = Point)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  facet_wrap(~variable, ncol = 2, scales = "free_y") +
  labs(x = "Horizon (quarters)", y = NULL,
       caption = "Shaded region: 68% Uhlig max-share bootstrap confidence band")
```

