---
title: "Replication: Galí (1999)"
subtitle: "Technology, Employment, and the Business Cycle: Do Technology Shocks Explain Aggregate Fluctuations?"
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

## Overview

This document replicates the main empirical results of @galí1999, which identifies technology shocks using the Blanchard-Quah (BQ) long-run restriction: only technology shocks can have a permanent effect on labor productivity. The VAR is estimated on US data in first differences, and level responses are recovered by cumulating the IRFs.

## Setup

```{r}
#| label: setup
library(tidyverse)
library(tidyMacro)
library(tictoc)

set_theme(ftheme_tidyMacro())
```

## Data

```{r}
#| label: data
data("Gali1999")

y <- Gali1999 |> select(Productivity, Hours) |> as.matrix()
dates <- Gali1999 |> pull(Date)

T <- nrow(y)
N <- ncol(y)
```

## VAR Estimation

```{r}
#| label: var-estimation

# Lag length selection via AIC
p <- fAICBIC(y, pmax = 4, c = 1)$aic
c <- 1

# Estimate VAR: y_t = c + A_1 y_{t-1} + ... + A_p y_{t-p} + e_t
var_result <- fVAR(y, p = p, c = c)

# Variance-covariance matrix of reduced-form residuals
Sigma <- var_result$sigma_full
```

## BQ Identification

Galí (1999) assumes only technology shocks have a permanent effect on labour productivity. This imposes a lower-triangular structure on the long-run multiplier matrix $C_1 K$, identified via:

$$
C_1 = \sum_{h=0}^{H} \Psi_h, \quad D_1 = \text{chol}(C_1 \Sigma C_1')^\top, \quad K = C_1^{-1} D_1
$$

```{r}
#| label: bq-identification

horizon <- 40
wold <- fwoldIRF(var_result, horizon = horizon)

# Long-run multiplier matrix
C1 <- apply(wold, c(1, 2), sum)

# Lower Cholesky of long-run covariance
D1 <- t(chol(C1 %*% Sigma %*% t(C1)))

# Structural impact matrix
K <- solve(C1, D1)

# BQ IRFs (in first differences)
irf_bq <- fbqIRF(wold, K)
```

## Bootstrap Confidence Bands

```{r}
#| label: bootstrap

tic()
boot_bq <- fbootstrapBQ(
  y = y,
  var_result = var_result,
  nboot = 1000,
  horizon = 40,
  prc = 68,
  bootscheme = "residual",
  cumulate = c(1, 2)
)
toc()
```

## Impulse Response Functions

Level responses are recovered by cumulating the first-differenced IRFs. The first shock is the technology shock; the second is the non-technology shock.

```{r}
#| label: fig-irf
#| fig-cap: "Impulse response functions to technology and non-technology shocks (68% confidence bands)"

source(here::here('R', 'fplotirf_bq.R'))
fplotirf_bq(
  point = irf_bq,
  upper = boot_bq$upper,
  lower = boot_bq$lower,
  varnames = c("LABPROD", "HOURS"),
  shocknames = c("Shock:Technology", "Shock:Non-Technology"),
  cumulate = c(1, 2),
  shocks = c(1, 2),
  prc = 68
)
```

## Forecast Error Variance Decomposition

```{r}
#| label: fig-fevd
#| fig-cap: "Forecast error variance decomposition"

# Cumulate BQ IRF along horizon (dim 3) to get level responses
irf_cumu <- irf_bq
for (h in 2:dim(irf_bq)[3]) {
  irf_cumu[,, h] <- irf_cumu[,, h] + irf_cumu[,, h - 1]
}

# FEVD — reuses Cholesky FEVD function (same maths)
fevd <- fevd_chol(irf_cumu, shock = 0)$fevd
fevd <- fevd[, c(2, 1), ] # optional, for vis purposes

fplot_vardec(
  fevd = fevd,
  varnames = c("LABPROD", "HOURS"),
  shocknames = c("Technology", "Non-Technology")
) +
  scale_fill_manual(values = tidyMacro_colors[1:2])
```

## Historical Decomposition

```{r}
#| label: fig-histdec
#| fig-cap: "Historical decomposition: contribution of technology shocks"

histdec_all <- vector("list", N)
ystar_all <- matrix(0, nrow = T - p, ncol = N)

for (series in seq_len(N)) {
  res <- fhistdec(y, var_result, K, series)
  histdec_all[[series]] <- res$histdec
  ystar_all[, series] <- res$ystar
}

names(histdec_all) <- c("LABPROD", "HOURS")

fplot_histdec(
  histdec_list = histdec_all,
  shock = 1,
  shockname = "Technology",
  dates = dates,
  p = p,
  facet_ncol = 1
) +
  scale_x_date(date_breaks = "6 years", date_labels = "%Y")
```

## Conditional Correlations

Following Galí (1999), we decompose the unconditional correlation between hours and productivity into shock-specific conditional correlations. The conditional correlation for shock $i$ is:

$$
\rho_i = \frac{\sum_{h=0}^{H} \Psi_{1i,h} \, \Psi_{2i,h}}{\sqrt{\left(\sum_{h=0}^{H} \Psi_{1i,h}^2\right)\left(\sum_{h=0}^{H} \Psi_{2i,h}^2\right)}}
$$

```{r}
#| label: cond-corr

# Dimensions of irf_bq: [variables, shocks, horizon]
N_shocks <- dim(irf_bq)[2]

# Numerator: for each shock i, sum of cross-products of IRFs of var 1 and var 2
numerator <- numeric(N_shocks)
for (i in 1:N_shocks) {
  temp         <- irf_bq[1, i, ] * irf_bq[2, i, ]
  numerator[i] <- sum(temp)
}

# Denominator: sqrt of product of sum-of-squares for each shock
irf_squared <- irf_bq^2
denominator <- numeric(N_shocks)
for (i in 1:N_shocks) {
  v1             <- sum(irf_squared[1, i, ])
  v2             <- sum(irf_squared[2, i, ])
  denominator[i] <- sqrt(v1 * v2)
}

# Conditional correlations (one per shock)
cond_corr <- numerator / denominator

# Unconditional correlations
uncond_corr <- cor(y)

cat("Unconditional correlation (Productivity, Hours):", round(uncond_corr[1, 2], 3), "\n")
cat("Conditional correlation | Technology shock:     ", round(cond_corr[1], 3), "\n")
cat("Conditional correlation | Non-Technology shock: ", round(cond_corr[2], 3), "\n")
```

The unconditional correlation between hours and productivity is small, at odds with standard RBC models. However, the conditional correlation driven by the technology shock is **negative**: a positive technology shock reduces hours worked. The non-technology shock yields a positive conditional correlation — the pattern RBC theory incorrectly attributed to technology.

