---
title: "Replication: Kaenzig (2021)"
subtitle: "The Macroeconomic Effects of Oil Supply News: Evidence from OPEC Announcements"
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

## Overview

This document replicates the main empirical results of @känzig2021 , which identifies oil supply news shocks using a high-frequency external instrument (proxy SVAR). The instrument captures exogenous variation in oil supply driven by OPEC production decisions, measured as oil futures price changes in a narrow window around OPEC announcements.

------------------------------------------------------------------------

## Setup

```{r}
#| label: setup
library(tidyverse)
library(tidyMacro)
library(tictoc)

set_theme(ftheme_tidyMacro())
```

------------------------------------------------------------------------

## Data

```{r}
#| label: data
data("Kaenzig2021")
finaldata <- Kaenzig2021 |>
    select(Oil_Price, World_Oil_Prod, World_Oil_Inven, World_IP, US_IP, US_CPI) |>
    as.matrix()

iv_oil <- Kaenzig2021 |>
    select(iv_kanzig_final) |>
    drop_na() |>
    as.matrix()

T <- nrow(finaldata)
N <- ncol(finaldata)

varnames  <- c("Oil Price", "World Oil Prod.", "World Oil Inven.",
               "World IP", "US IP", "US CPI")
shockname <- "Oil Shock"
```

------------------------------------------------------------------------

## VAR Estimation

```{r}
#| label: var-estimation

p <- 12
c <- 1

var_result <- fVAR(finaldata, p, c)
residuals  <- var_result$residuals
sigma_full <- var_result$sigma_full

# Wold IRFs
hor  <- 48
wold <- fwoldIRF(var_result, horizon = hor)
```

------------------------------------------------------------------------

## Proxy SVAR: MBB Bootstrap

```{r}
#| label: bootstrap-mbb

tic()
result_mbb <- fbootstrapIV_mbb(
    y          = finaldata,
    var_result = var_result,
    Z          = iv_oil,
    nboot      = 1000,
    blocksize  = 0,          # auto block size
    adjustZ    = c(1, 417),
    adjustu    = c(100, 516),
    policyvar  = 1,
    horizon    = hor,
    prc        = 90
)
toc()
```

------------------------------------------------------------------------

## Impulse Response Functions

```{r}
#| label: fig-irf
#| fig-cap: "IRFs to oil supply news shock — MBB bootstrap (90% confidence bands)"

fplotirf_iv(
    point     = result_mbb$meanirf * 10,
    upper     = result_mbb$upper * 10,
    lower     = result_mbb$lower * 10,
    varnames  = varnames,
    shockname = shockname,
    prc       = 90,
    facet_ncol = 3
)
```

------------------------------------------------------------------------

## IV Identification (Manual Two-Stage)

The proxy SVAR is identified manually following the two-stage IV approach. The instrument $Z_t$ is regressed on the reduced-form residual of the policy variable (oil price) to obtain the fitted values $\hat{u}_{p,t}$, which are then used to identify the structural impact vector $s$.

```{r}
#| label: iv-identification

# Align residuals and instrument to overlapping sample
u_p     <- residuals[, 1]
u_p_fin <- u_p[(length(u_p) - length(iv_oil) + 1):length(u_p)] |> as.matrix()
u_q_fin <- residuals[
    (nrow(residuals) - length(iv_oil) + 1):nrow(residuals),
    2:ncol(residuals)
] |> as.matrix()

u   <- residuals[(nrow(residuals) - length(iv_oil) + 1):nrow(residuals), ]
T1  <- nrow(u)
sigma <- (t(u) %*% u) / (T1 - 1 - p - N * p)
S     <- t(chol(sigma))

# Stage 1: regress u_p on instrument
ols1 <- fOLS(y = u_p_fin, X = iv_oil, c = 0)
uhat <- ols1$fitted_partial

# Stage 2: regress other residuals on uhat to get s_q / s_p
sq_sp <- solve(t(uhat) %*% uhat) %*% t(uhat) %*% u_q_fin

# Structural impact vector (normalised to unit oil price response)
s <- c(1, sq_sp[1], sq_sp[2], sq_sp[3], sq_sp[4], sq_sp[5])

# Structural IRFs
ivirf <- matrix(0, nrow = N, ncol = hor + 1)
for (h in 1:(hor + 1)) {
    ivirf[, h] <- wold[, , h] %*% s
}
```

------------------------------------------------------------------------

## Forecast Error Variance Decomposition

```{r}
#| label: fig-fevd
#| fig-cap: "Forecast error variance decomposition: share explained by oil supply news shock"

fevd_result <- fevd_iv(s, S, wold, N, hor, sigma, u, T1 = T1, p = p)

fplot_vardec(
    fevd       = fevd_result$fevd_iv,
    varnames   = varnames,
    shocknames = shockname
)
```

------------------------------------------------------------------------

## References

::: {#refs}
:::
